---
title: "homework-03"
author: Jocelyn Magner
date: 5/20/2026
format:
  pdf:
    toc: true #  display a table of contents
execute:
  message: false # use this to make sure messages don't show up
  warning: false # use this to make sure warnings don't show up
---

# Set up
```{r}
library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(ggeffects)

# store personal data
personal_data <- read_csv(here("data" , "personaldataproject.csv"))

# store kelp data
temp_kelp <- read_csv(here("data", "temp-kelp.csv"))
```

# Problems
## Problem 1. Giant Kelp Fronds
### a. An Appropriate Test
Pearson and Spearman correlations could be used to determine strength of the relationship between temperature and giant kelp frond elongation rate assuming observations are independent.
Pearson's correlation also assumes a linear relationship exists and it is a parametric test meaning it assumes variables are normally distributed. Spearman's correlation assumes a monotonic relationship and it is a nonparametric test meaning normality is not an assumption.

### b. Visualizing Data
Showing the relationship between temperature and giant kelp frond elongation rate.

```{r}
# plot raw data points to observe potential linear relationship

# base layer ggplot
ggplot(data = temp_kelp,
       # assign data to axes
       aes(x = temp_c,
           y = kelp_elong))+
# first layer: add scatter points
  geom_point(size = 2,
             stroke = 0.5,
             fill = "dodgerblue3",
             shape = 21)+
  
# relabel axis with units
  labs(x = "Temperature (°C)",
       y = "Giant Kelp Frond Growth Rate (cm/day)")+
# change theme
  theme_light()+
  theme(panel.grid = element_blank())
# data appears fairly linear, proceed with diagnostic tests
```


### c. Check Assumptions and Run Test

```{r}
#| echo: true
#| fig-width: 8
#| fig-height: 8
# create object storing linear model data to perform diagnostic tests
temp_kelp_model <- lm(kelp_elong~temp_c,
   data = temp_kelp)

# diagnostic plots
par(mfrow = c(2,2))
plot(temp_kelp_model)
```

#### Assumption of Pearson's correlation

• Linear relationship between variables:

Looking at the geom point graph from 1b. the relationship appears linear.

• Variables are continuous: 

Temperature and length are both continuous variables (they can be measured to decimal places or fractions).

• Variables are normally distributed:

Looking at diagnostic QQ plot, the data appears normally distributed since the data follows the line closely only deviating strongly in the tails which is acceptable.

• Independent observations:

Similar to the abalone observations presented in lecture, I think in order to have independent kelp frond growth observations they would need to be grown in individual tanks with specific regulated temperatures. Since the study averages temperatures and doesn't specify tanks or growing conditions this may not be completely met, but we are going to work with it.


### Run a correlation test

```{r}
#| echo: true
# Run Pearson Correlation test
cor.test(temp_kelp$kelp_elong, # predicted/y value
         temp_kelp$temp_c,     # predictor/x value
         method = "pearson")
```


### d. Results Communication

To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, I used a Pearson's Correlation test.
The data appears to be normally distributed based on the QQ plot diagnostic test meaning this parametric test is fitting.

We found a moderate-to-strong negative relationship between giant kelp frond elongation rate (in cm/day) and
Temperature (in degrees Celsius) (Pearson's r = -0.69, t(30) = -5.19, p < 0.01, α = 0.05).

### e. Test Implications

Our tests suggest a moderately strong statistically significant negative relationship (r = -0.69, p< 0.01) between temperature and kelp growth, meaning that giant kelp frond elongations rates decrease as water temperatures rise. Based on this information, to protect giant kelp populations, I would suggest targeting conservation efforts in the warmest environments.

### f.

## Problem 2. Personal Data
### a.
### b.

## Problem 3. Affective Visualization
### a.
### b.
### c.
### d.
### e.

## Problem 4. Statistical Critique
### a.
### b.
### c.
### d.

