---
title: "homework-03"
format: PDF
author: Jocelyn Magner
date: 5/20/2026
---

# Set up
```{r}
library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(ggeffects)

# store personal data
personal_data <- read_csv(here( "data" , "personaldataproject.csv"))

# store kelp data
temp_kelp <- read_csv(here("data", "temp-kelp.csv"))
```

# Problems
## Problem 1. Giant Kelp Fronds
### a. An Appropriate Test
Pearson and Spearman correlations could be used to determine strength of the relationship between temperature and giant kelp frond elongation rate assuming observations are independent.
Pearson's correlation also assumes a linear relationship exists and it is a parametric test meaning it assumes variables are normally distributed. Spearman's correlation assumes a monotonic relationship and it is a nonparametric test meaning normality is not an assumption.

### b. Visualizing Data
```{r}
# plot raw data points to observe potential relationship

# base layer ggplot
ggplot(data = temp_kelp,
       # assign data to axes
       aes(x = temp_c,
           y = kelp_elong))+
# first layer: add points
  geom_point(color = "dodgerblue3")+
# relabel axis wiht units
  labs(x = "Temperature (°C)",
       y = "Giant Kelp Frond Growth Rate (cm/day)")+
#change theme
    theme_light()
# data appears fairly linear, proceed with diagnostic tests
```

```{r}
# create object storing linear model data to perform diagnostic tests
temp_kelp_model <- lm(temp_c~kelp_elong,
   data = temp_kelp)
```

```{r}
# base R residuals
par(mfrow = c(2,2))
plot(temp_kelp_model)

# observations appear fairly normal making it a good fit for Pearson Correlation tes
```

```{r}
ggpredict(
  ?
  
  
  temp_kelp_model, terms = "temp_c [14:21 by=1]") %>% 
  plot(show_data = TRUE)
```


### c. 
### Check Assumptions

Assumptions of linear regression
• Linear relationship expected: A linear relationship is logical. I would expect temperature to affect 
• Independent errors
• Homoscedasticity
• Normally Distributed Error

Assumption of Pearson's correlation
 linear relationships between response and predictors
• independent errors
• homoscedastic errors (assessed using residuals)
• normally distributed errors (assessed using residuals)

### Run a correlation test

```{r}

```


### d.
### e.
### f.

## Problem 2. Personal Data
### a.
### b.

## Problem 3. Affective Visualization
### a.
### b.
### c.
### d.
### e.

## Problem 4. Statistical Critique
### a.
### b.
### c.
### d.

