---
title: "homework-03"
format: PDF
author: Jocelyn Magner
date: 5/20/2026
---

# Set up
```{r}
library(tidyverse)
library(here)
library(janitor)
library(readxl)

# store personal data
personal_data <- read_csv(here( "data" , "personaldataproject.csv"))

# store kelp data
temp_kelp <- read_csv(here("data", "temp-kelp.csv"))
```

# Problems
## Problem 1. Giant Kelp Fronds
### a. 
Pearson and Spearman correlations could be used to determine strength of the relationship between temperature and giant kelp frond elongation rate assuming observations are independent.
Pearson's correlation also assumes a linear relationship exists and it is a parametric test meaning it assumes variables are normally distributed. Spearman's correlation assumes a monotonic relationship and it is a nonparametric test meaning normality is not an assumption.

### b. Visualizing Data

```{r}
ggplot(data = )
```

### c.
### d.
### e.
### f.

## Problem 2. Personal Data
### a.
### b.

## Problem 3. Affective Visualization
### a.
### b.
### c.
### d.
### e.

## Problem 4. Statistical Critique
### a.
### b.
### c.
### d.

